Predictive modeling with cross validation

### 2017-08-24 model fitting/cross_validation.Rmd
assessed residuals + PH assumption

#results of stepAIC (including contain_answered even though it was dropped)
results <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday, data = train1)
summary(results)
AIC(results, k = 2) # 64969.63
```

```{r}
results2 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + 
    contain_unanswered + title_questionmark + sqrt(avg_tag_score) + 
    text_all_lower + update + prior_effort + ampm + weekday + 
    pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, 
    df = 0, method = "AIC"), data = train1)
AIC(results2, k = 2) # 64866.93

@@

```{r, residuals, echo = FALSE}
#=============================================
#martingale residuals
train1$mart <- residuals(model_full, type = "martingale")

ggplot(train1, aes(x = device_length, y = mart)) + 
  geom_point() + 
  geom_smooth() + 
  ggtitle(paste("Martingale Residuals for device_length")) + 
  scale_x_continuous("Device_length")

#=============================================
#deviance residuals 
train1$deviance <- residuals(model_full, type = "deviance")
ggplot(train1, aes(x = 1:nrow(train1), y = deviance)) + 
  geom_point() + 
  scale_x_continuous("Question index") + 
  scale_y_continuous(breaks = seq(-4, 4, by = 0.5)) + 
  ggtitle("Deviance Residuals")
# quite a few questions have a deviance residual of larger than 2.5

#=============================================
#schoenfeld residuals
schoen <- as.data.frame(residuals(model_full, type = "schoenfeld"))
# complete event times sorted from shortest to longest
schoen$comp_times <- sort(train1[train1$answered != 0,]$time_until_answer)

#=============================================
# score residuals
score <- as.data.frame(residuals(model_full, type = "score"))
# function to plot score residuals
plot_score <- function(column) {
  ggplot(score, aes(x = 1:nrow(score), y = score[[column]])) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Score residuals for", column)) + 
    scale_y_continuous("Score Residuals") + 
    scale_x_continuous("Question index")
}
columns <- names(score)
purrr::map(columns, plot_score)

#=============================================
# working with rms::which.influence to id influential points
c <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, x = TRUE, y = TRUE, data = train1)
influence <- rms::which.influence(c)
rms::show.influence(influence, train1)

#=============================================
# formal test of PH assumption
options(scipen=999)
ph_test <- data.frame(predictors = rownames(cox.zph(model_full)$table), cox.zph(model_full)$table) 
rownames(ph_test) = NULL
ph_test <- ph_test %>%
  arrange(p)

#=============================================
# addression PH assumption violations

# stratifying on new_category
strat_model <- coxph(Surv(time_until_answer, answered) ~ strata(new_category) + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm, data = train1)
AIC(strat_model, k = 2) # 49766.8
summary(strat_model)

# stratifying on new_user
strat_model1 <- coxph(Surv(time_until_answer, answered) ~ new_category + strata(new_user) + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(strat_model1, k = 2) # 60543.75
summary(strat_model1)

# stratifying on ampm 
strat_model2 <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + strata(ampm) + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(strat_model2, k = 2) # 54447.8
summary(strat_model2)

# stratifying on ampm, new_user, new_category
strat_model3 <- coxph(Surv(time_until_answer, answered) ~ strata(new_category) + strata(new_user) + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + strata(ampm) + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(strat_model3, k = 2) # 35370.81
summary(strat_model3)
```

@@

##### restricted cubic spline on text_length (had the smallest AIC)

* predictions on test data not working

```{r}
#  cv_rcstext <- function(train, test) {
# 
#   # building model on training data set
#   model <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(text_length, unname(quantile(train1$text_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train)
# 
#   metric <- cph(Surv(time_until_answer, answered) ~ p, data = train)
#   train_metrics
# 
#   # performance metrics for training data
#   train[["predictions"]] <- exp(predict(model, type = "lp"))
#   metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train)
#   train_metrics <- data.frame(HR = exp(metric$coefficients),
#                               LR = round(metric$stats[3],2),
#                               pval = round(metric$stats[5],2),
#                               Dxy = round(metric$stats[9],2),
#                               AIC = stats::AIC(metric, k = 2),
#                               Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train)$concordance)
# 
#   # predicting on test data
#   test[["predictions"]] <- exp(predict(model, newdata = test, type = "lp"))
# 
#   # computing performance metrics
#   metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
#   test_metrics <- data.frame(HR = exp(metric1$coefficients),
#                              LR = round(metric1$stats[3],2),
#                              pval = round(metric1$stats[5],2),
#                              Dxy = round(metric1$stats[9],2),
#                              AIC = stats::AIC(metric1, k = 2),
#                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)
# 
#   # returns data frame with train/test metrics
#   statistics <- rbind(train_metrics, test_metrics)
#   rownames(statistics) <- c("Training Data", "Test Data")
#   return(statistics)
# }
# 
# rcstext_models <- purrr::map2(trains, tests, ~cv_rcstext(.x, .y))
# splinenew_models

# Fitting model to the full data set

```{r, echo = FALSE}
x <- oshitar::variable_setup(x)

final <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(text_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = x)
summary(final)
AIC(final, k = 2) # 83051.61

x$predictions <- predict(final, type = "risk")
(performance <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = x))
AIC(performance, k = 2)

# extracting predicted survival probabilities
prob <- pec::predictSurvProb(final, newdata = x, times = c(0.5, 1, 1.5, 2, quantile(x$time_until_answer)))

#==================================================================================
# function to take in a data set, fit the model, output performance metrics, and allow you to specify times to predict survival probabilities

@@

>>>>

#results of stepAIC (including contain_answered even though it was dropped)
results <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday, data = train1)
summary(results)
AIC(results, k = 2) # 64969.63

# results of second stepAIC
results2 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + 
    contain_unanswered + title_questionmark + sqrt(avg_tag_score) + 
    text_all_lower + update + prior_effort + ampm + weekday + 
    pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, 
    df = 0, method = "AIC"), data = train1)
AIC(results2, k = 2) # 64866.93

final_model <- results2
```

@@

@@

```{r, echo = FALSE}
# function to take in a data set, fit the model, output performance metrics, and allow you to specify times to predict survival probabilities

@@

```{r, residuals, echo = FALSE}
# variables : new_category + as.factor(new_user) + 
#     contain_unanswered + title_questionmark + sqrt(avg_tag_score) + 
#     text_all_lower + update + prior_effort + ampm + weekday + 
#     pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, 
#     df = 0, method = "AIC") 
#=============================================
#martingale residuals
train1$mart <- residuals(final_model, type = "martingale")
train1$sqrt_score <- sqrt(train1$avg_tag_score)

ggplot(train1, aes(x = sqrt_score, y = mart)) + 
  geom_point() + 
  geom_smooth() + 
  ggtitle(paste("Martingale Residuals for device_length")) + 
  scale_x_continuous("Device_length")
# what do I assess for spline variables?

#=============================================
#deviance residuals 
train1$deviance <- residuals(final_model, type = "deviance")
ggplot(train1, aes(x = 1:nrow(train1), y = deviance)) + 
  geom_point() + 
  scale_x_continuous("Question index") + 
  scale_y_continuous(breaks = seq(-4, 4, by = 0.5)) + 
  ggtitle("Deviance Residuals")
# quite a few questions have a deviance residual of larger than 2.5 (use ggplotly to identify)

#=============================================
# schoenfeld residuals
schoen <- as.data.frame(residuals(final_model, type = "schoenfeld"))
# complete event times sorted from shortest to longest
schoen$comp_times <- sort(train1[train1$answered != 0,]$time_until_answer)

# plots indicate that new_user, ampm and a few others may have violated ph assumption

#=============================================
# formal test of PH assumption
options(scipen=999)
ph_test <- data.frame(predictors = rownames(cox.zph(final_model)$table), cox.zph(final_model)$table) 
rownames(ph_test) = NULL
ph_test <- ph_test %>%
  arrange(p)
ph_test
#=============================================
# score residuals
score <- as.data.frame(residuals(final_model, type = "score"))

# function to plot score residuals
plot_score <- function(column) {
  ggplot(score, aes(x = 1:nrow(score), y = score[[column]])) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Score residuals for", column)) + 
    scale_y_continuous("Score Residuals") + 
    scale_x_continuous("Question index")
}
columns <- names(score)
purrr::map(columns, plot_score)
# score residuals show that there are quite a few points that may be influential (check with ggplotly)
```
